### 2019-10-24 manuscript/manuscript.Rmd

---
title: Towards Explainability of Machine Learning Models in Insurance Pricing
author:
  - name: Kevin Kuo
    email: kevin@kasa.ai
    affiliation: a
    footnote: Corresponding Author
  - name: Daniel Lupton
    email: dlupton@taylorandmulder.com
    affiliation: b
address:
  - code: a
    address: Kasa AI, 3040 78th Ave SE \#1271, Mercer Island, WA 98040, USA
  - code: b
    address: Taylor & Mulder, 10508 Rivers Bend Lane, Potomac, MD 20854, USA
abstract: |
  Machine learning methods have garnered increasing interest among actuaries in recent years. However, their adoption by practitioners has been limited, partly due to the lack of transparency of these methods, as compared to generalized linear models. In this paper, we discuss the need for model interpretability in property & casualty insurance ratemaking, propose a framework for explaining models, and present a case study to illustrate the framework.

journal: "[journal]"
date: "`r Sys.Date()`"
bibliography: explain-pricing.bib
#linenumbers: true
numbersections: true
# csl: elsevier-harvard.csl
output: 
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
layout: preprint, 3p, twocolumn, letterpaper, 12pt
---

# Introduction

Risk classification for property & casualty (P&C) insurance rating has traditionally been done with one-way, or univariate, analysis techniques. In recent years, many insurers have moved towards using generalized linear models (GLM), a multivariate predictive modeling technique, which addresses many shortcomings of univariate approaches, and is currently considered the gold standard in insurance risk classification. At the same time, machine learning (ML) techniques such as deep neural networks have gained popularity in many industries due to their superior predictive performance over linear models [@lecunDeepLearning2015]. In fact, there is a fast growing body of literature on applying ML to P&C reserving [@kuoDeepTriangleDeep2018; @wuthrichMachineLearning2018; @gabrielliNeuralNetwork2019a; @gabrielliNeuralNetwork2019]. However, these ML techniques, often considered to be completely “black box”, have been less successful in gaining adoption in pricing, which is a regulated discipline and requires a certain amount of transparency in models.

If insurers can gain more insight into how ML models behave in risk classification contexts, it would increase their ability to reassure regulators and the public that accepted ratemaking principles are met. Being able to charge more accurate premiums would, in turn, make the risk transfer system more efficient and contribute to the betterment of society. In this paper, we aim to take a step towards liberating actuaries from the confines of linear models in pricing projects, by proposing a framework for explaining ML models for ratemaking that regulators, practitioners, and researchers in actuarial science can build upon.

The rest of this paper is organized as follows: Section \ref{ratemaking} provides an overview of P&C ratemaking, Section \ref{need} discusses the importance of interpretation, Section \ref{interpretability} discusses model interpretability in the context of ratemaking and proposes specific tasks for model explanation, Section \ref{application} describes current model interpretation techniques and applies them to the tasks defined in the previous section, and Section \ref{conclusion} concludes.

# Property and Casualty Ratemaking {#ratemaking}

## History of Ratemaking

Early classification ratemaking procedures were typically univariate in nature. For example, @lange_1966 notes that (at that time) most major lines of insurance used univariate methods based around the same principle: distributing an overall indication to territorial relativities or classification relativities based on the extent to which they deviated from the average experience.

@bailey_simon_1960 introduced minimum bias methods, which were expanded throughout the 60s, 70s, and 80s. As computing power developed, minimum bias began to give away to GLMs, with papers such as @brown_1988 and @mildenhall_1999 bridging the gap between the methods.

Arguably, GLMs predate minimum bias procedures by a significant margin. The term was coined by @nelder_wedderburn_1972, but generalizations of least squares linear regression date back at least to the 1930s. Like minimum bias methods, GLMs did not become mainstream in actuarial science for some time. For example, the syllabus of basic education does not seem to include any mention of GLMs prior to @brown_1988 in the 1990 syllabus for basic education. From there, GLMs seem to have received only passing mention until 2006 with the introduction of @anderson_2005 to the syllabus.

## Machine Learning in Ratemaking

Paralelling the development of GLM was the development of machine learning algorithms throughout the middle part of the 20th century. Detailed histories of machine learning may be found in sources such as @nilsson_2009 and @wang_raj_2017. Consistent with GLMs, machine learning was relatively unpopular in actuarial science until the last ten years as computing power has become cheaper and more easily available and as machine learning software packages have obviated the need for developing analyses from scratch each time an analysis is performed. Due to the breadth of machine learning as a field, it is difficult to identify the first time it entered the CAS syllabus; however, cluster analysis (in the form of k-means) seems to have been first included in 2011 with @robertson_2009. More recently, the MAS-I and MAS-II exams introduced in 2018 have included machine learning explicitly.

Within the area of ratemaking, machine learning is still in its infancy. A significant portion of machine learning applications to ratemaking has been in the context of automobile telematics, such as @gao_2018, @gao_2018_2, @gao_2019, @roel_2018, or @wuthrich_2017. Presumably this focus has been a result of the high-dimensionality and complexity of telematics data, making it a field in which the unique abilities of machine learning techniques give a clear advantage over traditional approaches.

Outside of telematics, @yang_2018 uses a gradient tree-boosting approach to capture non-linearities that would be a challenge for GLMs. @henckaerts_2018 makes use of generalized additive models (GAM) to improve predictions of GLMs. Many researchers, in an apparent effort to demonstrate the range of possibilities and advantages of machine learning, have approached the topic by comparing many different machine learning algorithms within a single study, such as in @dugas_2003, @noll_2018, @spedicato_2018. These studies make use of such varied techniques as regression trees, boosting machines, support vector machines, and neural networks.

## Ratemaking Process

Regardless of the method employed for determining this risk of various classifications, the actual process of setting rate relativities typically involves some variation of the following steps:

1. Obtain relevant policy-level data
2. Prepare data for analysis
3. Perform analysis on the data, employing desired method or methods to estimate needed rate relativites
4. Select final rate relativities based on rate indications
5. Present rates to the reglator, including explanation of the steps followed to derive the rates
6. Answer questions from regulators regarding the method employed

The focus of this paper is on steps 5 and 6. In particular, rate regulators are concerned with whether rates are inadequate, excessive, or unfairly discriminatory. In many states, rate filings that exceed certain thresholds for magnitude of rate changes or filings that make use of new or sophisticated predictive models may be subject to intense regulatory scrutiny. In these cases, it is necessary to be able to explain the results of the modeling process in a way that is understandable without sacrificing statistical rigor.

It should be noted that communicating results is not simply a method of passing regulatory muster. Generating interpretable modeling output is an important - even essential - facet of model checking. Therefore, the techniques discussed in this paper may be viewed from the lens of providing useful information to regulators, but they should also be considered as part of a thorough vetting of any rating model.

# The Need to Understand the Black Box {#need}

Within the actuarial profession, Actuarial Standard of Practice 41 ("Actuarial Communications") notes that "...another actuary qualified in the same practice area [should be able to] make an objective appraisal of the reasonableness of the actuary's work as presented in the actuarial report." [@asop_41] Underlying this requirement is an assumption that the hypothetical other actuary qualified in the same practice area is adequately familiar with the relevant techniques employed. Although the syllabus of basic education is constantly changing, there has at times been an assumption that all techniques and assumptions that have ever been a part of the syllabus of basic education needn't be explained from first principles in general actuarial communications, and that an actuary practicing in the same field should be able to make an objective appraisal of the results from the methods found in the syllabus. [This can be supported by reference to the edits to ASOP 38 over time - originally it was designed to be about all models, but they revised it to be about only models that incorporate specialized knowledge outside of the actuary's area of expertise... do we need this citation, though?] This is notable because, beginning with the introduction of the MAS-I and MAS-II examination in July of 2018, several machine learning models were formally included in the syllabus of basic education. These exams cover a wide range of topics, such as splines, clustering algorithms, decision trees, boosting, and principle components analysis [@cas_syllabus_2018].

Nevertheless, machine learning poses something of a special challenge for ASOP 41 for several reasons. Machine learning models can be very ad hoc compared to traditional statistical models. Because many machine learning models are deterministic, they may not admit of standard metrics for model comparison (e.g., it's not straightforward to calculate an AIC over a neural network). In addition, machine learning methods are often combined into ensembles that may not be easily separated and that may, as a collection, cease to resemble a single standard version of a model. Complicating matters still further, machine learning models can be "black boxes" insofar as the final form of response curve cannot be easily predicted and may depend heavily on the available data (which may not, in turn, be available to the reviewer).

This last item raises a final interesting issue. GLMs and their ilk are often fitted using one of a handful of standard and well-understood approaches (e.g., maximum likelihood estimation). However, this is not possible in general with machine learning models, as machine learning algorithms often use loss surfaces that are very complex such that it may not be feasible to calculate the global minimum of the surface. Certainly, closed form representations of the loss surfaces are not generally available. For this reason, the training phase of a machine learning model is, in many ways, just as important to one's understanding as the model form and the data on which the model is fitted. Because the final model result is inseparable from these three components (training method, model form, and data), it is not generally adequate to just know the method employed to make an objective appraisal of the reasonableness of the model. More information is necessary.

Of course, these comments only apply to the actuarial profession. Outside of the actuarial profession, communication of results may be more challenging. A 2017 survey conducted by the Casualty Actuarial and Statistical Task Force of the National Association of Insurance Commissioners found that the plurality of responding regulators identified "Filing complexity and/or a lack of resources or expertise" as a key challenge that impedes their ability to review GLMs or other predictive models [@naic_summer_2017]. Given that machine learning algorithms are generally regarded as more complex than GLMs, this implies that the challenge of communicating machine learning model results is significant.

In response to the same survey, 33 state regulators noted that it would be helpful or very helpful for the NAIC to develop information and tools to assist in reviewing rate filings based on GLMs, and 34 noted that it would be helpful to develop similar items to assist in reviewing "Other Advanced Modeling Techniques." One outgrowth of this need was the development of a white paper, @naic_white_paper, on best practices for regulatory review of predictive models. The white paper focuses on review of GLMs, particularly with respect to private passenger automobile and homeowners' insurance. Some of the guidance offered in this regard is therefore not strictly applicable to the review of machine learning models. For example, as previously noted, p-values are not a concept that translates well to deterministic machine learning algorithms. However, among the guidance applicable to machine learning algorithms are the following:

* Determine the extent the model causes premium disruption for individual policyholders, and how the insurer will
explain the disruption to individual consumers that inquire about it.
* Determine that individual input characteristics to a predictive model are related to the expected loss or expense
differences in risk. Each input characteristic should have an intuitive or demonstrable actual relationship to expected loss or expense.
* Determine that individual outputs from a predictive model and their associated selected relativities are not unfairly discriminatory.

These items are by no means exhaustive, but they pertain to the concept of model interpretability for ratemaking that we develop next.

# Interpretability in the Ratemaking Context {#interpretability}

In this section, we attempt to develop a working definition of interpretability for ratemaking applications. While we will not provide a comprehensive survey of the prolific and fast evolving ML interpretability literature, we draw from it as appropriate in setting the stage for our discussion. Even among researchers in ML, there is not a consensus on the definition of interpretability; here are a few from frequently cited papers:

- Ability to explain or to present in understandable terms to a human [@doshi-velezRigorousScience2017];
- The degree to which an observer can understand the cause of a decision [@biranExplanationJustification2017]; and
- A method is interpretable if a user can correctly and efficiently predict the method’s results [@kimExamplesAre2016].

We motivate our discussion by considering several aspects of interpretability. As we proceed through the points below, we aim to arrive at a more scoped and relevant definition of what it means to attain interpretability in a pricing model. In the remainder of this section, we clarify a couple concepts regarding interpretable *classes* of models and and the computational transparency of ML models, outline findings on the social aspects of ML explanations most relevant to our use case, then discuss a potential framework for implementing ML interpretability in practice.

## Not All Linear Models are Interpretable {#linear-models}

Let us look at some examples of models that are considered by many to be interpretable. The GLM is probably the most oft-cited example of such a model (or rather, modeling framework) in the actuarial literature. Given a set of inputs, we can easily reason about what the output of the model is. As an illustrative example, consider a claim severity model with driver age, sex, and vehicle age as features; assuming a log link function and letting $Y$ denotes the response, we have

\begin{align}
& \log(E[Y]) = \beta_0 + \beta_1 \cdot \text{age} \nonumber \\
& \quad + \beta_2 \cdot \text{vehicle\_age} + \beta_3 \cdot \text{sex}_{\text{male}}.
\end{align}

Here, we can tell, for example, what the model would predict for the expected severity if we were to increase age by a certain amount, *all else being equal*, because the relationship between the feature and the response is simply multiplication by the coefficient $\beta_1$ and applying the inverse link function. Another commonly cited example of an interpretable model is a decision tree. An illustrative example is shown in Figure \@ref(fig:tree-plot1). Here, the prediction is arrived at by following a sequence of if-else decisions.

```{r tree-plot1, echo = FALSE, fig.cap="A simple decision tree for loss cost prediction.", out.width="100%", fig.align='center'}
knitr::include_graphics("figures/tree-plot1.png")
```

Now, it is worth pointing out that, when declaring that GLM or decision trees are interpretable models, we are implicitly assuming that we are considering only a handful  of features. In fact, the ease at which we can reason about a model rapidly declines as the number of features, transformations, and interactions increase, as in the following (somewhat pathological) example:

\begin{align}\label{eq:badglm}
& \log(E[Y]) = \beta_0 + \beta_1 \cdot \text{age} + \beta_2 \cdot \text{vehicle\_age} \nonumber\\
& \quad + \beta_3\cdot\text{vehicle\_age}^2 \nonumber \\
& \quad + \beta_4 \cdot \text{age} \cdot \text{vehicle\_age} \nonumber \\
& \quad + \beta_5 \cdot \text{sex}_{\text{male}} + \beta_6 \cdot \text{sex}_{\text{male}} \cdot \text{age}. 
\end{align}

Similarly, one can see that in Figure \@ref(fig:tree-plot2), large and deep trees are tough to reason about. In other words, even when working within the framework of an "interpretable" class of models, we may still end up with something that many would consider "black box."

```{r tree-plot2, echo = FALSE, fig.cap="A more complex decision tree. This is still much simpler than typical realistic examples.", out.width="60%", fig.align='center',  fig.env='figure*'}
knitr::include_graphics("figures/tree-plot2.png")
```

## The Machinery is Not a Secret

Another occasional misconception is that we have no visibility into how some ML models *compute* predictions, which render them uninterpretable. Outside of proprietary algorithms, all common ML models, including neural networks, gradient boosted trees, and random forests, are well studied and have large bodies of literature documenting their inner workings. As an example, a fitted neural network model on tabular data is simply a composition of linear transformations followed with nonlinear activation functions. As in Equation \ref{eq:badglm}, one can write down the mathematical equation for calculating the prediction given some inputs, but it may be difficult for a human to reason about it. We show later that we can still provide explanations of completely "black box" models, but is important to note that ML model predictions are still governed by mathematical rules.

## Explanations are Contextual

Proposed by @hiltonConversationalProcesses1990 and interpreted by @millerExplanationArtificial2017 in the context of ML, model explanations are conversations or social interactions. One consequence of this identification is that explanations need to be *relevant* to the *audience*. In developing, filing, and operationalizing a pricing model, one needs to accommodate a variety of stakeholders, each of whom has a different set of questions, assumptions, and technical capacity. First, there are internal stakeholders at the company, which includes management and underwriters. While some of the individuals in this audience may be technical, they are likely less familiar with predictive modeling techniques than the actuaries and data scientists who build the models. Next, we have the regulators, who may have limited resources to review the models, and will focus on a specific list of questions motivated by statute. Finally, we have the public, who has an interest (perhaps more so than the other parties) because their lives are directly affected by the premiums charged.

It is interesting to note that, the modelers, who are most familiar with the models, tend to be same people designing and communicating the explanations. This poses a challenge that @millerExplainableAI2017 calls "inmates running the asylum", where the modelers design explanations for *themselves* rather than the intended audience. For example, they may be interested in technical questions like extrapolation behavior and shape the explanations accordingly, which may be lost on the prospective policyholder.

Another point outlined in Miller's survey [@millerExplanationArtificial2017] is that explanations are *contrastive*. In other words, people are often interested in not why something happened, but rather why it happened instead of something else. For example, policyholders might not care exactly how their auto premiums are computed, but would like to know why they are being charged more than their coworkers who drive similar vehicles. As an extension, policyholders may want to know what they can change in order to obtain lower premiums.

## Asking and Answering the Right Questions

With the above considerations in mind, we propose a potential framework for interpreting ML models for insurance pricing: the actuarial profession, in collaboration with regulators and representatives of the public, define a set of questions to be answered by explanations accompanying ML models, along with acceptance criteria and examples of successful explanations. In other words, interpretability for our purposes is defined as the ability of a model's explanations to answer the posed questions.

These questions should relate to existing guidelines, such as those described in @naic_white_paper and outlined in Section \ref{need}, standards of practice, and regulation, and in fact should not be specific only to ML models. By standardizing on a set of questions, we reduce the burden of both companies and regulators; this is especially important for the latter, who are already resource constrained facing increasing variety of models being filed. This format should also be familiar to actuaries who are accustomed to adhering to specific guidelines in, for example, ASOPs. Like the ASOPs, We envision that these questions and guidelines will be continually updated to reflect feedback obtained and advances in research.

While the realization of a set of such guidelines is an ambitious undertaking beyond the scope of this paper, we present in the next section a sample set of questions and techniques one can leverage to answer them. The goal of these case studies is twofold: to more concretely illustrate the proposed framework, and to expose the actuarial audience to modern ML interpretation techniques.

# Applying Model Interpretation Techniques {#application}

Now that we have established a framework for model interpretation in the form of asking and answering relevant questions, we demonstrate examples of such exchanges via an illustrative case study. Analytically, our starting point is a fitted deep neural network model for predicting loss costs. As the modeling details are of secondary importance, they are available in Appendix \@ref{model-dev}. The questions that we ask of the model are as follows:

1. What are the most important predictors in the model?
2. How does the predicted loss cost change, on average, as we change an input?
3. For a particular policyholder, how does each characteristic contribute to the loss cost prediction?

In our discussion, we adopt the organization of techniques and some notation presented in @molnar2018interpretable and @biecekPMVEE, which are comprehensive references on the most established ML interpretation techniques.

## A Simplified View of Interpretation Techniques

Before we dive into the answering questions, we provide some background on model interpretation techniques. As the literature is quite sizeable and fast moving, we do not attempt a full taxomony of techniques. The goal is to orient ourselves among broad categories of techniques, so we can map them to tasks indicated by the questions being asked. Model interpretation techniques can be categorized across two dimensions: intrinsic vs. post-hoc and global vs. local.

### Intrinsic vs. Post-hoc

Intrinsic model interpretation draws conclusions from the structure of the fitted model and are what we typically associate with "interpretable" classes of models. This is only viable with models with simple structures, such as the sparse linear model and shallow decision tree we see in Section \ref{linear-models}, where we arrive at explanations by reading off parameter estimates or a few decision rules. For algorithms that produce models with complex structure that do not lend themselves easily to intrinsic exploration, we can appeal to post-hoc techniques. This class of techniques interrogate the model by presenting it with data for and observing the prediction behavior of the model. These techniques are concerned with only the inputs and outputs, and hence are *agnostic* of the model itself, which means they can also be applied to simple models. Since most useful ML models have a level of complexity beyond the threshold of intrinsic interpretability, we focus on model-agnostic techniques in our case study. As we will see later on, the data that we present to the models are usually some perturbed variations of test data.

### Global vs. Local

Along the other dimension, we categorize model interpretations as global, or model-level, and local, or instance-level. The former class provides insights with respect to the model as a whole. Some examples of these eplanations include variable importances and sensitivities, on average, of the predicted response with respect to individual predictors. In our case study, questions 1 and 2 are associated with this category. On the other hand, question 3 pertains to an individual prediction, which would fall in the local interpretation category. In addition to individual variable attribution, we can also inquire about what would happen to the current predicted response if we were to perturb specific predictor variables.

## Answering the Questions

Having aligned the questions with the categories of interpretation techniques, we now introduce a selection of appropriate techniques to answer them.

### Variable Importance

"What are the most important predictors in the model?"

For linear models and their generalizations, and some ML models, measures of variable importance can be obtained from the fitted model structure. In the case of GLMs, one can observe the magnitudes of the $t$-statistics, whereas for random forests, one can use out-of-bag errors [@breimanRandomForests2001]. For more complex models, such as the neural network in our case study, we need to devise another approach.

We follow the methodology of permutation feature importance as described in @fisherAllModels2018, and utilize the notation introduced by @biecekPMVEE. The gist of the technique is as follows: to see how important a variable is, we make predictions without it and see how much worse off we are in terms of accuracy. One way to achieve this would be to re-fit the model many times (as many times as the number of variables.) However, this may be intractable with lengthy model training times or large number of variables, so a more popular approach is to instead keep the same fitted model but permute the values of each predictor.

More formally, let $y$ denote the vector of responses, $X$ denote the matrix of predictor variables, $\widehat{f}$ denote the fitted model, and $L = \mathcal{L}(\widehat{f}(X), y)$ denote the value of the loss function, which is mean squared error in the case of regression. Now, if $\widetilde{X}^{j}$ denotes the predictor matrix where the $j$th variable has been permutated, then we can compute the loss with the permuted dataset as $L^{-j} = \mathcal{L}(\widehat{f}(\widetilde{X}^{j}), y)$. With this, we define the absolute variable importance $VI^{j}$ as $L^{-j} - L$ or the relative variable importance as $VI_{R}^{j} = \frac{L^{-j}}{L}$.

In Figure \@ref(fig:fi), we show a plot of variable importances. In our particular example, we see that the "make" variable contributes most to the accuracy of the model with "sex" contributing the least. Note that these measures do not provide information regarding the directional sensitivity of the predictors on the response. This provides a way for the audience to quickly glance at the most relevant variables, and ask further question as necessary. For example, if a variable that can potentially be a proxy for a discriminatory variable appears as very important, more investigation would be warranted.

```{r fi, out.width='100%', fig.align = "center", fig.cap="Permutation feature importances for ML model.", echo = FALSE}
knitr::include_graphics("figures/fi-plot.png")
```

### Partial Dependence Plots

"How does the predicted loss cost change, on average, as we change an input?"

For this question, we again consider first how it would be answered in the GLM setting. When the input predictor in question is continuous, we can answer the question by looking at the estimated coefficient, which provides the change in the response per unit change in the predictor (on the scale of the linear predictor). For non-parametric models and neural networks, where no coefficients are available, we can appeal to partial dependence plots (PDP), first proposed by @friedmanGreedyFunction2001 for gradient boosting machines (GBM).

```{r pdp, out.width='70%', fig.align = "center", fig.cap="Partial dependence plot for ML model.", echo = FALSE,  fig.env='figure*', fig.pos='h'}
knitr::include_graphics("figures/pdp-plot.png")
```

```{r breakdown, out.width='80%', fig.align = "center", fig.cap="Variable contribution plot for ML model.", echo = FALSE, fig.env = 'figure*', fig.pos = 'h'}
knitr::include_graphics("figures/breakdown-plot.png")
```

To describe PDP, we need to introduce some additional notation. Let $x^j$ denote the input variable of interest. Then we define the partial dependence function as

\begin{equation}
h(z) = E_{X^{-j}}[\widehat{f}(x|x^j = z)],
\end{equation}

where the expectation is taken over the distribution of the other predictor variables. In other words, we marginalize them out so we can focus on the relationship between the predicted response and the variable of interest. Empirically, we estimate $h$ by

\begin{equation}
\widehat{h}(z) = \frac{1}{N}\sum_{i = 1}^{N}\widehat{f}(x_i|x_i^{j} = z),
\end{equation}

where $N$ is the number of records in our dataset.

In Figure \@ref(fig:pdp), we exhibit the PDP for the "vehicle age" variable. We see that the average predicted loss cost decreases with vehicle age until the latter is around 18. Note that the ticks on the $x$-axis tell us that the data is quite thin for vehicle age greater than 18, so the apparent upward trend to the right is driven by just a few data points. With this in mind, we can conclude that as cars get older, we expect the loss cost predictions generated by our model to decrease.

The question posed here is particularly important for regulators, who would like to know whether each variable affects the prediction in the direction that is expected, based on intuition, experience, and existing models. During the model development stage, PDP can also be used as a reasonableness test for candidate models by identifying unexpected relationships for the analyst to investigate.

### Variable Attribution

"For a particular policyholder, how does each characteristic contribute to the loss cost prediction?"

In the previous two examples, we look at model-level explanations; now we move on to one where we investigate one particular prediction instance. As before, we consider how we would approach the question for linear models. For a GLM with a log link common in ratemaking applications, for example, we simply start with the base rate, then the exponentiated coefficients would have multiplicative effects on the final rate. Similar to the previous examples, for ML models in general we do not have directly interpretable weights. Instead, one way to arrive at variable contributions is calculating the change in the expected model prediction for each predictor, conditioned on other predictors.

Formally, for a fitted model $\widehat{f}$ and a specific instance $x_*$, we would like to decompose the model prediction $\widehat{f}(x_*)$ into

\begin{equation}
\widehat{f}(x_* ) = v_0 + \sum_{j=1}^p v(j, x_*),
\end{equation}

where $v(j, x_*)$ denotes the contribution of the $j$th variable in instance $x_*$, defined as

\begin{align}\label{eq:var-contrib}
& v(j, x_* ) = E_X[\widehat{f}(X) | X^1 = x_* ^1, \dots, X^j = x_* ^j] \nonumber\\
& \quad - E_X[\widehat{f}(X^1 = x_* ^1, \dots, X^{j-1} = x_*^{j - 1})].
\end{align}

Hence, the contribution of the $j$th variable to the prediction is the incremental change in the expected model prediction when we set $X^j = x_*^ j$ assuming the other variables take their values in $x_*$. Note here that this definition implies that the order in which we consider the variables affects the results. Empirically, the expectations in \@ref(eq:var-contrib) are calculated by sampling the test dataset.

In Figure \@ref(fig:breakdown) we exhibit a waterfall plot of variable contributions. The "intercept" value denotes the average model prediction. The predicted loss cost for this particular policyholder is slightly less than average; the characteristics that makes this policyholder more risky are the fact that he is a male between the ages of 18 and 25; counteracting the risky driver characteristics are the vehicle properties: it is a GM vehicle built domestically and is seven years old.

Instance-level explanations are useful for investigating specific problematic predictions generated by the model. Regulators and model reviewers may be interested in variable contributions for the safest and riskiest policyholders to see if they conform to intuition. A policyholder with a particularly high premium may wish to find out what of their characteristics contribute to it, and may follow up with a question about how he can lower it, which would require another type of explanation.

## Other Techniques

(TODO: briefly discuss other stuff no covered)

# Conclusion {#conclusion}

(TODO)

References {#references .unnumbered}
==========
<div id="refs"></div>

\appendix

Appendix {.unnumbered}
=============

# ML Model Development {#model-dev}

In this appendix, we describe the ML model and the data used to train it. Note that, for our paper, the ultimate goal of the modeling procedure is to develop something that can produce predictions. As a result, we do not follow standard practices for supervised learning model development. However, for the sake of completeness and reproducibility, we include the details of the process here.

## Data

We use data from the AUTOSEG ("Automobile Statistics System") of Brazil's Superintendence of Private Insurance (SUSEP). The organization maintains policy-characteristics-level data, including claim counts and amounts, for all insured automobiles in Brazil. The data contains variables from policyholder characteristics to losses by peril. (TODO: finish)

## Model

(TODO)